### 2014-03-07 src/methodology.Rnw
update

One of our research questions was to compare the utility of operator
selection --- especially statement deletion --- with random sampling of mutants,
Since we use PIT, which is a bytecode oriented mutation engine, implementing statement deletion
directly is very hard. However, we can approximate statement deletion by combining different
bytecode operators, as given below.

The basic statement deletion was mdeled on the work by Deng et al.~\cite{deng2013empirical}. The operations
on single statements was modeled using \textit{VOID-M-CALLS,NONV-M-CALLS,CONSTR-CALLS,E-MEMBER-VAR,RM-INCR} for simple statements, and using \textit{RM-COND} for control structures, which replaced the boolean condition with false, resulting in removal of the block of statements. The operator for return values was modeled using \textit{RETURN-VALS}, which is similar.
The operators for \textit{while,for}, and \textit{if} statements were modeled using \textit{DELETE-COND}, which replaced the boolean condition with true, which removed the effect of conditional. The \textit{switch} statement deletion was modeled using \textit{RM-SWITCH} which replaced the first 100 labels with default label, resulting in the switch element getting deleted. (Due to the constraints of the architecture of PIT only the first 100 labels were replaced). The \textit{try catch} was not necessary at bytecode level.
Finally, to get an accurate estimation of the effect of true statement deletion, we grouped the mutants by line, and considered a single virtual mutant (That is, the total number of mutants is equal to the total number of lines). Further, killing any mutant from a line resulted in marking the virtual mutant for that line as killed. This gave us the mutation score for the virtual statement deletion operator.
We note that the approximation of simple statement deletion, especially when arithmetic operators are involved is not complete. However, as seen in Figure~\ref{fig:op}, the number of mutants produced by \textit{MATH} operator is very small. We also note that our approximation does not account for the increase in ease of detection when multiple mutations are combined together due to coupling effect. This also means that not all the lines may be mutated since there may be no applicable operators. However, given the constraints of a bytecode based mutation system, we believe that our procedure is reasonable.

@@

While the sampling criteria is simple, it is not immediately clear when we are faced with program elements that produced too few mutants for a particular fraction to yield at-least one mutant. For example, for a line that produced 15 mutants, the number of mutants that can be chosen in \textit{(x\%, line, 1/64)} is less than 1. We analyzed the results of the following criterias: 1) to round the resulting number to the nearest integer, and choose that value as sampling (in effect, ignoring elements that are less complex), 2) to apply \textit{ceil} function to the resulting value, ensuring that there is atleast one mutant per program element, and 3) to perform the sampling in such a way that it chooses a mutant with with the fractional probability (this is the method followed by Zhang et al.~\cite{zhang2013ase}).

@@

>>>>

One of our research questions was to compare the utility of operator
selection --- especially statement deletion --- with random sampling of mutants,
Since we use PIT, which is a bytecode oriented mutation engine, implementing statement deletion
directly is very hard. However, we can approximate statement deletion by combining different
bytecode operators, as explained later.

@@

While the sampling criteria is simple, it is not immediately clear when we are faced with program elements that produced too few mutants for a particular fraction to yield at-least one mutant. For example, for a line that produced 15 mutants, the number of mutants that can be chosen in \textit{(x\%, line, 1/64)} is less than 1. We analyzed the results of the following criterias: 1) to round the resulting number to the nearest integer, and choose that value as sampling (in effect, ignoring elements that are less complex), 2) to apply \textit{ceil} function to the resulting value, ensuring that there is atleast one mutant per program element, and 3) to perform the sampling in such a way that it chooses a mutant with with the fractional probability (this is the method followed by Zhang et al.~\cite{zhang2013ase}).
% TODO: If the criterias ceil and round does not make sense, we might remove them to gain space.

@@

\textit{Statement Deletion:}
The basic statement deletion was modeled on the work by Deng et
al.~\cite{deng2013empirical}. The operations on single statements was modeled
using \textit{VOID-M-CALLS,NONV-M-CALLS,CONSTR-CALLS,E-MEMBER-VAR,RM-INCR}
for simple statements, and using \textit{RM-COND} for control structures,
which replaced the boolean condition with false, resulting in removal of
the block of statements. The operator for return values was modeled using
\textit{RETURN-VALS}, which is similar.  The operators for \textit{while,for},
and \textit{if} statements were modeled using \textit{DELETE-COND},
which replaced the boolean condition with true, which removed the effect
of conditional. The \textit{switch} statement deletion was modeled using
\textit{RM-SWITCH} which replaced the first 100 labels with default label,
resulting in the switch element getting deleted. (Due to the constraints
of the architecture of PIT only the first 100 labels were replaced). The
\textit{try catch} was not necessary at bytecode level.  Finally, to get an
accurate estimation of the effect of true statement deletion, we grouped
the mutants by line, and considered a single virtual mutant (That is, the
total number of mutants is equal to the total number of lines). Further,
killing any mutant from a line resulted in marking the virtual mutant
for that line as killed. This gave us the mutation score for the virtual
statement deletion operator.  We note that the approximation of simple
statement deletion, especially when arithmetic operators are involved is not
complete. However, as seen in Figure~\ref{fig:op}, the number of mutants
produced by \textit{MATH} operator is very small. We also note that our
approximation does not account for the increase in ease of detection when
multiple mutations are combined together due to coupling effect. This also
means that not all the lines may be mutated since there may be no applicable
operators. However, given the constraints of a bytecode based mutation system,
we believe that our procedure is reasonable.
